---
title: "Commuting between OA, 2011"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=TRUE, results= 'hide', message=FALSE}
library(igraph)
library(knitr)
# library(corrplot)
# library(corrgram)
library(rgdal)
library(tidyverse)
#library(spdplyr)
# library(geojsonio)
library(stplanr)
library(ggplot2)
#library(leaflet)
#library(SpatialPosition)
library(stargazer)
library(rprojroot)
library(cleangeo)
library(tmap)

# This is the project path
path <- find_rstudio_root_file()
```

## Commuting flows

```{r}
path.data <- paste0(path, "/data/wf02ew_oa_v1.csv" )

commuting <- read_csv(path.data, 
                      col_names = F) %>% 
  rename(o = X1,
         d = X2,
         weight = X3) %>%              # commuting flows
  filter(substr(d, 1, 1) == "E" |      # Keep England, 
           substr(o, 1, 1) == "E") %>% # drop Wales, Scotland, NI, and 
                                       # non-geographical OD codes 
                                       # OD0000001 = Mainly work at or from home
                                       # OD0000002 = Offshore installation
                                       # OD0000003 = No fixed place
                                       # OD0000004 = Outside UK
  glimpse()
```

The below is with to exclude the intra-OA flows. 
**NOT TO RUN FOR NOW*

```{r eval=FALSE}
commuting.intra <- commuting %>%
  dplyr::filter(`Area of usual residence` == `Area of workplace`)
#commuting <- commuting %>%
#  dplyr::filter(`Area of usual residence` != `Area of workplace`)
```

Given the size of the network I am using the `fast and greedy` community detection algorithm. 
The optimal number of communities is $11$.
The problem is that when I try to force $communities$ = $constituencies$ the communities after the $11th$ have only $1$ $OA$.
Therefore, I present the optimal solution here.


```{r communities}
net <-graph_from_data_frame(commuting, directed = TRUE, vertices = NULL)

# I need to look on the sensitivity of the below
# net.un <- as.undirected(net,
#                         mode=c("mutual"),
#                         edge.attr.comb = igraph_opt("edge.attr.comb"))

net.un <- as.undirected(net,
                        mode="collapse",
                        edge.attr.comb = igraph_opt("edge.attr.comb"))

# fast and greedy
fg <- net.un %>% 
  fastgreedy.community() 

V(net.un)$group <- membership(fg) # assign membership to vertices

sizes(fg)

# fg$membership <- cut_at(fg, 543) 
# sizes(fg)
# for cut_at 543, the first 11 communities have > 1

names <- vertex_attr(net.un)[1]

df <- cbind(as.data.frame(names),
            fg$membership) %>% 
  rename('community' = 'fg$membership')

path.out <- paste0(path, "/outputs/fg.csv")
write_csv(df, path.out)
```

Other community detection attempts.

**Not to run**

```{r eval=FALSE}
# not hierarchical, I can't define cut_at
# l <- net.un %>% 
#   cluster_louvain() %>% 
#   cut_at(no = 543)



# Use spinglass to create a set number of communities
sg <- net %>% 
  cluster_spinglass(spins = 543) # 543 = the  number of constituencies allocated
                                 # to England for the 2023 Review

# Use hierarchical methods and cut_at to create a set number of communities 
walk <- net %>% 
  cluster_walktrap() %>% cut_at(no = 543) 

eb <- net %>% cluster_edge_betweenness() %>% cut_at(no = 10)


membership(net.l)
length(fg)
modularity(net.l)
print(net.l)
sizes(net.l)
is_hierarchical(net.l)

plot_dendrogram(net.l)

plot(lc, g)

```

## Mapping the communities

```{r mapping, include=TRUE, results= 'markup', message=FALSE}

# OA from https://geoportal.statistics.gov.uk/datasets/ons::output-areas-december-2011-boundaries-ew-bgc-1/about

path.oa <- paste0(path, "/data/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")

oa <- readOGR(path.oa)

head(oa@data)

# Let's try to map our output. We need the local authorities shape file we have 
# already loaded

# Before we do any further analysis we make sure to convert the la object to WGS84 
# Longitude / Latitude Coordinate System.
wgs84 = '+proj=longlat +datum=WGS84'
oa <- spTransform(oa, CRS(wgs84) )

# This is the @data element of the la spatial object
head(oa@data)

# read data in case the community detection is not run
df <- read_csv(path.out)

# Let's merge it with the communities data frame
oa@data <- merge(oa@data, df, by.x = "OA11CD", by.y = "name", all.x = T)
sum(is.na(oa@data$community))
oa@data$community <- as.factor(oa@data$community)

# clean ploygons
oa <- clgeo_Clean(oa)

path.map <- paste0(path, "/outputs/fg_commuting.png")

fg.commuting <- tm_shape(oa) + 
  tm_fill("community", 
          title= "Communities (fast and greedy)") + 
  tm_layout("Commuting flows 2011")

tmap_save(fg.commuting, path.map)
```

