---
title: "Historical migration"
output:
  html_document:
    toc: true
    toc_float: true
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../outputs/")
  })
---

```{r setup, include=TRUE, results= 'hide', message=FALSE}
library(igraph)
library(knitr)
# library(corrplot)
# library(corrgram)
library(rgdal)
library(tidyverse)
#library(spdplyr)
# library(geojsonio)
library(stplanr)
library(ggplot2)
#library(leaflet)
#library(SpatialPosition)
library(stargazer)
library(rprojroot)
library(cleangeo)
library(tmap)

# This is the project path
path <- find_rstudio_root_file()
```

## MIgration flows

```{r}
path.data <- paste0(path, "/data/LEFT_HOME_1911_XY.txt" )

migration <- read_csv(path.data, col_names = T) %>% 
  mutate(o = paste0(BP_X, "_", BP_Y ),    # unique origins coded as X_Y
         d = paste0(RES_X, "_", RES_Y),   # unique destinations coded as X_Y
         od = paste0(o, "-", d)) %>%      # unique od pairs coded as X_Y-X_Y (o_o_d_d)
  group_by(od) %>%          
  summarise(weight = n()) %>%             # flows for each od pair 
  #migration %>% 
  separate(od, c("o","d"), sep = "-") %>% # recreate unique o and d
  glimpse()                               # 12099 unique d and 
                                          # 21913 unique o
# length(unique(migration$o))

test <- migration %>% 
  filter(o == d)
```

The below is with to exclude the intra-OA flows. 
**NOT TO RUN FOR NOW*

```{r eval=FALSE}

migration.intra <- migration %>% 
  filter(o == d)                   # 6849 / 3104633

# migration <- migration %>% 
#   filter(o != d)
```

Given the size of the network I am using the `fast and greedy` community detection algorithm. 
The optimal number of communities is $14$.

The problem is that when I try to force $communities$ = $constituencies$ the communities after the $11th$ have only $1$ $OA$.
Therefore, I present the optimal solution here.


```{r communities}
net.migration <-graph_from_data_frame(migration, directed = TRUE, vertices = NULL)

# I need to look on the sensitivity of the below
# net.un <- as.undirected(net,
#                         mode=c("mutual"),
#                         edge.attr.comb = igraph_opt("edge.attr.comb"))

net.migration.un <- as.undirected(net.migration,
                        mode="collapse",
                        edge.attr.comb = igraph_opt("edge.attr.comb"))

# fast and greedy
fg.migration <- net.migration.un %>% 
  fastgreedy.community() 

V(net.migration.un)$group <- membership(fg.migration) # assign membership to vertices

sizes(fg.migration)

# fg$membership <- cut_at(fg, 543) 
# sizes(fg)
# for cut_at 543, the first 11 communities have > 1

names.migration <- vertex_attr(net.migration.un)[1]

df.migration <- cbind(as.data.frame(names.migration),
            fg.migration$membership) %>% 
  rename('community' = 'fg.migration$membership')

path.out <- paste0(path, "/outputs/fg_migration.csv")
write_csv(df.migration, path.out)
```

Other community detection attempts.

**Not to run**

```{r eval=FALSE}
# not hierarchical, I can't define cut_at
# l <- net.un %>% 
#   cluster_louvain() %>% 
#   cut_at(no = 543)



# Use spinglass to create a set number of communities
sg <- net %>% 
  cluster_spinglass(spins = 543) # 543 = the  number of constituencies allocated
                                 # to England for the 2023 Review

# Use hierarchical methods and cut_at to create a set number of communities 
walk <- net %>% 
  cluster_walktrap() %>% cut_at(no = 543) 

eb <- net %>% cluster_edge_betweenness() %>% cut_at(no = 10)


membership(net.l)
length(fg)
modularity(net.l)
print(net.l)
sizes(net.l)
is_hierarchical(net.l)

plot_dendrogram(net.l)

plot(lc, g)

```

## Mapping the communities

```{r mapping, include=TRUE, results= 'asis', message=FALSE}

# OA from https://geoportal.statistics.gov.uk/datasets/ons::output-areas-december-2011-boundaries-ew-bgc-1/about

path.oa <- paste0(path, "/data/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")

oa <- readOGR(path.oa)

head(oa@data)

# Let's try to map our output. We need the local authorities shape file we have 
# already loaded

# Before we do any further analysis we make sure to convert the la object to WGS84 
# Longitude / Latitude Coordinate System.
wgs84 = '+proj=longlat +datum=WGS84'
oa <- spTransform(oa, CRS(wgs84) )

# This is the @data element of the la spatial object
head(oa@data)

# read data in case the community detection is not run
df <- read_csv(path.out)

# Let's merge it with the communities data frame
oa@data <- merge(oa@data, df, by.x = "OA11CD", by.y = "name", all.x = T)
sum(is.na(oa@data$community))
oa@data$community <- as.factor(oa@data$community)

# clean ploygons
oa <- clgeo_Clean(oa)

path.map <- paste0(path, "/outputs/fg_commuting.png")

fg.commuting <- tm_shape(oa) + 
  tm_fill("community", 
          title= "Communities (fast and greedy)") + 
  tm_layout("Commuting flows 2011")

tmap_save(fg.commuting, path.map)
```

